---
title: "Evaporation Pan Rate Calculation"
subtitle: "with weather factors"
author: "Yolanda Ming, Manish Venumuddula, and Anne Fetrow"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output:
  html_document: 
    df_print: paged # omit to disable paged data table output
    css: stylesheet.css # omit if no need for custom stylesheet
    number_sections: yes # change to no for unnumbered sections
    toc: yes # change to no to disable table of contents
    toc_float: true # change to false to keep toc at the top
    toc_depth: 3 # change to specify which headings to include in toc
    code_folding: show # change to hide to hide code by default
editor_options:
  chunk_output_type: console # change to inline to show output inline
---

# Setup
Load libraries and read in custom function scripts.
```{r setup, echo = TRUE, message=FALSE, warning=FALSE}
# Load necessary libraries
library(readr)        # For reading CSV files
library(readxl)       # For reading Excel files
library(lubridate)    # For handling date-time data
library(dplyr)        # For data manipulation
library(ggplot2)      # For plotting
library(RColorBrewer) # For plotting
library(tidyverse) 
library(latex2exp)
library(egg)
library(viridis)
library(plotly)

# source all relevant scripting files
source(file.path("scripts", "plotting_functions.R"))
source(file.path("scripts", "table_functions.R"))
source(file.path("scripts", "eva_climateFactor_functions.R"))
source(file.path("scripts", "eva_calc_functions.R"))

# global knitting options for automatic saving of all plots as .png and .pdf
knitr::opts_chunk$set(
  dev = c("png", "pdf"), fig.keep = "all",
  dev.args = list(pdf = list(encoding = "WinAnsi", useDingbats = FALSE)),
  fig.path = file.path("plots", paste0(gsub("\\.[Rr]md", "", knitr::current_input()), "_"))  )
```

Define working directories
```{r}
# Define paths for input data and output plots
# ACF filepath
data_directory <- "/Users/annefetrow/Library/CloudStorage/GoogleDrive-fetrowa@gmail.com/Shared drives/Mono Lake Hydrology Project/Supplies & Equipment/Evaporation Pan/Data processing/Mono-Evap-Pan/data/eva pan"
# Yolanda filepath
# data_directory <- "C:/Users/24468/Desktop/Research/SEAS-HYDRO/Mono Lake/Mono-Evap-Pan/data/eva pan"
plot_directory <- file.path(data_directory, "plot")

# Ensure the plot directory exists
if (!dir.exists(plot_directory)) {
  dir.create(plot_directory)
}
```

Specify theme information for plots produced below.
```{r theme}
# Set color palette
coul <- brewer.pal(5, "Set2")

theme_figure <- function(legend = TRUE, grid = TRUE, plot_margin = c(1, 1, 1, 1), 
                         text_size = 14, axis_text_size = NULL, axis_x_rotate = 0) { # changed font size from 12 to 18 for AGU
  the_theme <- theme_bw() +
    theme(text = element_text(size = text_size),
          plot.background = element_blank(), panel.background = element_blank(),
          panel.border = element_rect(color="black", size=1),
          strip.background = element_rect(color="black", linetype = 1),
          plot.margin = unit(plot_margin, "mm")#, 
          # legend.position = c(1, 1), # 1 refers to the top-right corner
          # legend.justification = c(1, 1) # Justify the legend to the top-right corner
      )
  # adjust grid
  if(!grid)
    the_theme <- the_theme + theme(panel.grid = element_blank())
  else
    the_theme <- the_theme + theme(panel.grid.minor = element_blank())
  # adjust legend
  if (!legend)
    the_theme <- the_theme + theme(legend.position = "none")
  # overwrite axis text size if provided
  if (!is.null(axis_text_size))
    the_theme <- the_theme +
      theme(axis.text = element_text(size = axis_text_size))
  # axis rotation
  if (axis_x_rotate != 0) {
    the_theme <- the_theme +
      theme(axis.text.x = element_text(angle = axis_x_rotate, vjust = 0.5, hjust = 1))
  }
  return(the_theme)
} 

# add to plots for consistent them
#   + theme_figure()

#for large scale poster figure output
  #           theme(text = element_text(size = 24),
  #           plot.background = element_blank(), panel.background = element_blank(),
  #           panel.border = element_rect(color="black", size=1),
  #           strip.background = element_rect(color="black", linetype = 1),
  #          # plot.margin = unit(plot_margin=1, "mm"), 
  #           legend.position = c(0.195, 0.90), # 1 refers to the top-right corner, c(0.195, 0.98),
  #           legend.justification = c(1, 1), # Justify the legend to the top-right corner
  #           legend.title = element_text(size = 24), # Adjust legend title font size
  #           legend.text = element_text(size = 20), # Adjust legend text font size
  #           legend.margin = margin(t = 5, r = 5, b = 5, l = 10), # Adjust legend margin
  #           legend.box.background = element_rect(color = "black", size = 1) # Add a black box around the legend
  #               )
```

# Data
```{r}
# Import and preprocess water level data from multiple sources
data1 <- water_level_data_import(file.path(data_directory, 'ML_EP_20230907.csv'))
data2 <- water_level_data_import(file.path(data_directory, '2023-11-2MLevap (2).csv'))
h <- c(data1$h, data2$h)
date_time <- as.POSIXct(c(data1$date_time, data2$date_time))
```

# Evap Rate 
```{r}
# Calculate evaporation rate
eva_data <- eva_rate_calc(h, date_time)
eva_rate <- eva_data$eva_rate
eva_date_time <- as.POSIXct(eva_data$eva_date_time)
eva_data$date_only <- as.Date(eva_data$eva_date_time) # Convert the eva_date_time column to Date objects to remove the time component

# Compute average evaporation rate
eva_rate_avg <- mean(eva_rate[eva_rate < 0.75], na.rm = TRUE) # Ignore evaporation that is larger than 0.75 mm/hr
gg_daily <- daily_eva_rate_calc(eva_data)
gg_monthly <- monthly_eva_rate_clac(eva_data)

# Create data frames for plotting evaporation and precipitation
df_eva <- data.frame(eva_date_time, eva_rate)
df_precip <- precip(data_directory, "Mono Lake_Evaporation_Factors_Daily_Yolanda.xlsx")
```

# Plots
```{r}
# Generate and save plot for evaporation rate and precipitation
eva_prep_plot <- ggplot() +
  geom_bar(data = df_eva, aes(x = eva_date_time, y = eva_rate), stat = "identity", fill = "blue", color = "blue") +
  geom_bar(data = df_precip, aes(x = date_time, y = P_d), stat = "identity", fill = "red", color = "red") +
  geom_hline(yintercept = eva_rate_avg, linetype = "dashed", color = "blue") +
  labs(x = "Date Time", y = "mm/hr", title = "Evaporation Rate and Precipitation") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_x_datetime(date_breaks = "2 weeks", date_labels = "%b %Y")  # Adjust date_breaks as needed
eva_prep_plot
ggplotly(eva_prep_plot)

# Save plot to file
ggsave(filename = "evarate_prep_barplot.pdf", plot = eva_prep_plot, path = plot_directory, device = "pdf", width = 9, height = 6)
ggsave(filename = "daily_evap_rate.pdf", plot = gg_daily, path = plot_directory, device = "pdf", width = 9, height = 6)
ggsave(filename = "monthly_evap_rate.pdf", plot = gg_monthly, path = plot_directory, device = "pdf", width = 9, height = 6)

# Plot weather factors and save to PDF
weather_factors_filename <- "Mono Lake_Evaporation_Factors_Daily_Yolanda.xlsx"
plot_weather_factors(file.path(data_directory, weather_factors_filename), plot_directory)
```

# ACF comments
You've done really super work! 

I took your code and put it into an Rmarkdown (.Rmd) file type to make running the code chunks easier and plot viewing more immediate. Additionally, this way, we can knit this code into .html that can be hosted online. I moved the code from "evap_rate_with_weather_factor.R" to this file and created markdown headers.

My largest overall comment is that in order to make this code as transparent as possible, I think it would be best bring in the functions you've written to calculate various terms (ex: evap rate) into one comprehensive code. This likely breaks a rule of true "coders", but we know this code will ultimately be used by beginner to intermediate experience users and it will decrease the amount of "chasing" someone would have to do to understand the hidden math in the functions. Essentially, I think it would be best to combine this Rmarkdown and the "R_from_MARLAB.R" file. The R_from_MARLAB.R" code is where the majority of the important math is happening and the functions defined in there can be added to this code.

Additionally for each step in the functions you've written, it would be good if you could add a comment about what each line is doing so someone could follow along. Think of this as essentially what you would say or write on a board if you were explaining the function to someone. 
Also, we ultimately want this one Rmarkdown to workout of the same working directory and currently the functions being in the R_from_MATLAB.R file has specific filepaths in it that would have to be changed by each user -- ideally these functions wouldn't require a specific filepath but rather reference a dataframe that is specified in this Rmarkdown file by the user. Happy to talk about this in person, b/c this is tricky to explain.

Next steps:
- Consider the suggestions above and see if its possible to create a more unified code that reads in the necessary data, defines the calculation functions, does the actual evap rate calculations, and generates plots into one Rmarkdown.
- Write an explanation of how you've calculated evaporation rate including information about how precipitation periods are excluded and include equations and variable definitions. The vision here is that this could be included in a methods section of a manuscript for how evaporation rate calculations were generated. We can also then break that explanation up and embedd it in this code to explain the steps to future users. The "Method/Equation Summary" googledoc you shared already has a lot fo what you'll need, but add paragraphs to explain the approach around the equations. 
- Write up figure captions to accompany all the plots that detail the various components (ex: what do the red bars mean vs blue in the evaporation rate bar plot). Format this as if you were going to embed it in a report or manuscript
