---
title: "L2S_Comparison"
author: "Yolanda Ming"
date: "`r Sys.Date()`"
output: html_document
---
# Overview
This code analyzes and visualizes monthly evaporation rates for Mono Lake using multiple estimation methods and model predictions. The analysis integrates observed pan evaporation data, theoretical evaporation calculations from the Penman method, and the Webb method, alongside L2S hydrological model outputs. The code outputs a PNG `l2s_comparison.png` to `output` folder.

# Initialization

Before running the code, We defined some global settings and load necessary libraries. You don't need to change the settings, but please make sure you have all the libraries downloaded to your computer. 

```{r, echo=TRUE, message=FALSE, warning=FALSE, results='hide'}
# Initialize
cat("\014")
rm(list = ls())
graphics.off()

# Load necessary libraries
library(dplyr)
library(readr)
library(lubridate)
library(ggplot2)
library(vroom)
library(gridExtra)
library(grid)
library(cowplot)
```

You will need to change the working directory to your own personal computer's directory when you download it.

```{r, echo=TRUE, message=FALSE, warning=FALSE, results='hide'}
# Define a global base directory, please change it when download the code to your personal desktop
global_work_dir <- "C:/Users/24468/Desktop/Research/SEAS-HYDRO/Mono Lake/Mono-Evap-Pan"
global_save_dir <- file.path(global_work_dir, "plots")
```


# Function for Image Output

Here, we define a help function to shorten the main body of the code. You don't need to change anything for the code to run. However, there are some parameters that determine the output plot format and size, etc. Feel free to change that based on your liking.

```{r, echo=TRUE, message=FALSE, warning=FALSE, results='hide'}
export_plot_to_png <- function(plot, file_name, save_dir = global_save_dir, width = 1200, base_height = 200, res = 150, num_rows = 1, combined = FALSE) {
  
  # Adjust height only if it's a combined plot and there are multiple rows
  if (combined && num_rows > 1) {
    # Add extra space for the x-axis labels on the first plot (top plot)
    extra_space_for_labels <- 300  # Adjust this value as needed
  } else {
    extra_space_for_labels <- 0  # No extra space needed for non-combined or single-row plots
    base_height = 500
  }
  
  # Calculate the total height
  height <- base_height * num_rows + extra_space_for_labels
  
  # Create the directory if it doesn't exist
  dir.create(save_dir, showWarnings = FALSE, recursive = TRUE)
  
  # Construct the full file path
  file_path <- file.path(save_dir, file_name)
  
  # Open a PNG device with calculated height
  png(filename = file_path, width = width, height = height, res = res)
  
  # Check if the input is a single plot or a grid object
  if (inherits(plot, "ggplot")) {
    # For a single ggplot, use print
    print(plot)
  } else {
    # For grid-arranged objects, use grid.draw
    grid.draw(plot)
  }
  
  # Close the device
  dev.off()
  
  cat("Plot exported to:", file_path, "\n")
}
```

# Load L2S Data

The code below loads data from l2s model results (credit to UM SEAS HYDRO Lab's researcher Anna Gossard, Manish Venumuddula, Alexander VanDeWeghe1, Caleb Wegener, and
professor Andrew D. Gronewold). The model predictions contain mean value, upper and lower boundaries.

```{r, echo=TRUE, message=FALSE, warning=FALSE, results='hide'}
# Load the CSV file
file_path <- file.path(global_work_dir, "data", "l2s", "Joker_L2SWBM_jSum_Formatted.csv")
data <- read.csv(file_path)

# Convert the 'Date' column to Date type
data$Date <- as.Date(data$Date, format = "%Y-%m-%d")

# Extract the relevant columns (evaporation rate, upper boundary, lower boundary)
l2s_evaporation_rate <- 304.8 / 30.44 * data[, 17]
l2s_upper_boundary <- 304.8 / 30.44 * data[, 19]
l2s_lower_boundary <- 304.8 / 30.44 * data[, 23]

# Create a new data frame with the necessary columns
tbl <- data.frame(Date = data$Date, l2s_evaporation_rate, l2s_upper_boundary, l2s_lower_boundary)

# Extract the month from the Date column
tbl$Month <- month(tbl$Date)

# Calculate the monthly averages across all years using dplyr
l2s_monthly_avg <- tbl %>%
  group_by(Month) %>%
  summarise(
    monthly_avg_evaporation_rate = mean(l2s_evaporation_rate, na.rm = TRUE),
    monthly_avg_upper_boundary = mean(l2s_upper_boundary, na.rm = TRUE),
    monthly_avg_lower_boundary = mean(l2s_lower_boundary, na.rm = TRUE)
  )

# Create Date column for the monthly averages (using an arbitrary year, e.g., 2000)
l2s_monthly_avg$Date <- as.Date(paste(2000, l2s_monthly_avg$Month, 1, sep = "-"))

# View the resulting data frame
print(l2s_monthly_avg)
```

# Load Converted Evaporation Estimate

We also loaded pan evaporation data and evaporation estimates from both Webb and Penman methods.

```{r, echo=TRUE, message=FALSE, warning=FALSE, results='hide'}
# Load the CSV file with stringsAsFactors = FALSE to avoid factor conversion
file_path <- file.path(global_work_dir, "output", "eva_estimate_Penman_RStudio.csv")
data <- read.csv(file_path, stringsAsFactors = FALSE)

# Check the first few rows of the Date column to ensure it's in the correct format
head(data$Date)

# Convert the 'Date' column to Date type (YYYY-MM-DD format)
data$Date <- as.Date(data$Date, format = "%Y-%m-%d")  # YYYY-MM-DD format

# Extract the relevant columns
dates <- data$Date
lake_from_pan_eva <- data$Lake_From_Pan_Eva_mm_d
lake_from_theoretical_eva <- data$Lake_From_Theoretical_Eva_mm_d
pan_eva <- data$Pan_Eva_mm_d

# Create a data frame with the necessary columns
tbl <- data.frame(dates, lake_from_pan_eva, lake_from_theoretical_eva, pan_eva)

# Extract the year and month from the dates
tbl$Year <- year(tbl$dates)
tbl$Month <- month(tbl$dates)

# Calculate the monthly averages across all years using dplyr
monthly_avg <- tbl %>%
  group_by(Month) %>%
  summarise(
    monthly_avg_lake_from_pan_eva = mean(lake_from_pan_eva, na.rm = TRUE),
    monthly_avg_lake_from_theoretical_eva = mean(lake_from_theoretical_eva, na.rm = TRUE),
    monthly_avg_pan_eva = mean(pan_eva, na.rm = TRUE)
  )

# Create Date column for the monthly averages (using an arbitrary year, e.g., 2000)
monthly_avg$YearMonth <- as.Date(paste(2000, monthly_avg$Month, 1, sep = "-"))

# View the resulting data frame
print(monthly_avg)


# Load the new CSV file
file_path <- file.path(global_work_dir, "output", "eva_coeff_Webb_RStudio.csv")
data_webb <- read.csv(file_path, stringsAsFactors = FALSE)

# Convert the 'Date' column to Date type
data_webb$Date <- ymd(data_webb$Date)

# Extract Year-Month as a string (e.g., "2024-July")
data_webb$YearMonth <- paste(year(data_webb$Date), month(data_webb$Date, label = TRUE, abbr = FALSE), sep = "-")

# Replace the year part with "2000" while keeping the month intact
data_webb$YearMonth <- gsub("^\\d{4}", "2000", data_webb$YearMonth)

# Convert to Date (assuming 'YearMonth' is now in "2000-July" format)
data_webb$YearMonth <- as.Date(paste0(data_webb$YearMonth, "-01"), format = "%Y-%B-%d")

# Sort the table by YearMonth
data_webb <- data_webb %>% arrange(YearMonth)

# Calculate the monthly averages for the 'Coeff' column
monthly_avg_webb <- data_webb %>%
  group_by(YearMonth) %>%
  summarise(mean_Coeff = mean(Coeff, na.rm = TRUE))

# Combine the two data frames to get the result
monthly_avg <- left_join(monthly_avg, monthly_avg_webb, by = "YearMonth")

# Add the new result as a column (monthly evaporation based on coefficients)
monthly_avg$monthly_avg_lake_from_webb <- monthly_avg$monthly_avg_pan_eva * monthly_avg_webb$mean_Coeff
```

# Visualize Comparison Between Model and Evaporation Estimate

We plotted the l2s model as line plot as they contains detailed daily model results. Evaporation estimate, along with original pan evaporation data, is plotted as monthly average using dots.

```{r, echo=TRUE, message=FALSE, warning=FALSE, results='hide'}
# Start ggplot without specifying data initially
p <- ggplot()

# Plot the l2s data as lines
p <- p +
  # Evaporation Rate (Solid Black Line)
  geom_line(data = l2s_monthly_avg, 
            aes(x = Date, y = monthly_avg_evaporation_rate, color = "Evaporation Rate"), 
            size = 1) +
  
  # Upper Boundary (Dashed Black Line)
  geom_line(data = l2s_monthly_avg, 
            aes(x = Date, y = monthly_avg_upper_boundary, color = "Upper Boundary"), 
            linetype = "dashed", size = 1) +
  
  # Lower Boundary (Dashed Black Line)
  geom_line(data = l2s_monthly_avg, 
            aes(x = Date, y = monthly_avg_lower_boundary, color = "Lower Boundary"), 
            linetype = "dashed", size = 1) +
  
  # Plot the Penman Method Scatter Points
  # Lake Theoretical Eva, Penman (Red Filled Circles)
  geom_point(data = monthly_avg, 
             aes(x = YearMonth, y = monthly_avg_lake_from_theoretical_eva, color = "Penman, Theoretical"), 
             size = 3, shape = 21, fill = "red") +
  
  # Actual Pan Eva (Green Filled Circles)
  geom_point(data = monthly_avg, 
             aes(x = YearMonth, y = monthly_avg_pan_eva, color = "Actual Pan Eva"), 
             size = 3, shape = 21, fill = "green") +
  
  # Plot the Webb Method Scatter Points
  # Lake Eva, Webb (Blue Filled Circles)
  geom_point(data = monthly_avg, 
             aes(x = YearMonth, y = monthly_avg_lake_from_webb, color = "Webb"), 
             size = 3, shape = 21, fill = "blue") +
  
  # Lake Eva Incorporating Pan Eva, Penman (Yellow Filled Circles)
  geom_point(data = monthly_avg, 
             aes(x = YearMonth, y = monthly_avg_lake_from_pan_eva, color = "Penman, Pan Eva"), 
             size = 3, shape = 21, fill = "yellow") +
  
  # Labels and Title
  labs(
    y = 'Evaporation Rate (mm/d)',
    x = 'Month',
    title = 'Monthly Average Lake Evaporation Rate from Multiple Methods with l2s Model',
    color = 'Legend'
  ) +
  
  # Customize X-axis to show each month with labels
  scale_x_date(
    date_breaks = "1 month", 
    date_labels = "%B",
    limits = c(as.Date("2000-01-01"), as.Date("2000-12-01"))
  ) +
  
  # Define manual colors for the legend
  scale_color_manual(
    values = c(
      "Evaporation Rate" = "black",
      "Upper Boundary" = "black",
      "Lower Boundary" = "black",
      "Penman, Theoretical" = "red",
      "Actual Pan Eva" = "green",
      "Webb" = "blue",
      "Penman, Pan Eva" = "yellow"
    )
  ) +
  
  # Apply a minimal theme and rotate x-axis labels for better readability
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5),
    legend.position = "top"  # Change "none" to "top" or another position
  )
  
  # Add grid lines
  theme(panel.grid.major = element_line(color = "grey80"),
        panel.grid.minor = element_line(color = "grey90"))

# Print the plot
print(p)

export_plot_to_png(p, file_name = "l2s_comparison.png", num_rows = 1, combined = FALSE)
```

